### 2021-07-15 src/paper/report.tex
2762

At the core of our agent-based model \citep[][we review more literature in Supplementary
Material~\ref{sec:literature_review}]{Aleta2020,Hinch2020} are physical contacts between
heterogeneous agents (Figure~\ref{fig:model_contacts_infections}). Each contact between
an infectious individual and somebody susceptible to the disease bears the risk of
transmitting the virus. Contacts occur in up to four networks: Within the household, at
work, at school, or in other settings (leisure activities, grocery shopping, medical
appointments, etc.). Some contacts recur regularly, others occur at random. Empirical
applications can take the population and household structure from census data and the
network-specific frequencies of contacts from diary data measuring contacts before the
pandemic \citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, meeting frequencies
depend on age and geographical location (see Supplementary
Material~\ref{subsec:assortativity}).

The four contact networks are chosen so that the most common NPIs can be modeled in
great detail. NPIs affect the number of contacts or the risk of transmitting the disease
upon having physical contact. The effect of different NPIs will generally vary across
contact types. For example, a mandate to work from
home will reduce the number of work contacts to zero for a fraction of the working
population. Schools and daycare can be closed entirely, operate at reduced
capacity---including an alternating schedule---, or implement mitigation measures like
masking requirements or air filters \citep{Lessler2021}. In any setting, measures like
masking requirements would reduce the probability of infection
\citep{Cheng2021}.

@@

In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age. A
possible infection progresses as shown in Figure~\ref{fig:model_disease_progression}. We
differentiate between an initial period of infection without being infectious or showing
symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring
intensive care, and recovery or death \citep[similar to][]{Aleta2020}. The probabilities
of transitioning between these states depend on age; their duration is random and
calibrated to medical literature (see Supplementary
Material~\ref{sub:data_course_of_disease}). Conditional on the type of contact,
infectiousness is independent of age \citep{Jones2021}.

The model includes several other features, which are crucial to describe the evolution of
the pandemic in 2020-2021. New virus strains with different infectiousness profiles may
appear. Vaccines may become available. During the vaccine roll-out, priority may depend
on age and occupation; vaccine hesitancy is modelled by some individuals refusing
vaccination offers. With some probability, vaccinated agents become immune and do not
transmit the virus \citep{Hunter2021, LevineTiefenbrun2021, Petter2021, Pritchard2021}.

We include two types of tests. Polymerase chain reaction (PCR) tests reveal whether an
individual is infected or not; there is no uncertainty to the result. PCR tests require
at least one day to be processed and there are aggregate capacity constraints. In
contrast, rapid antigen tests yield immediate results. Specificity and sensitivity of
these tests is set according to data analyzed in \cite{Bruemmer2021, Smith2021};
sensitivity depends on the timing of the test relative to the onset of infectiousness.
Figure~\ref{fig:model_rapid_tests} shows our model for rapid test demand. Schools may
require staff and students to be tested regularly. Rapid tests may be offered by
employers to on-site workers. Individuals may demand tests for private reasons, which
include having plans to meet other people, showing symptoms of CoViD-19, and a household
member having tested positively for the virus. We endow each agent with an individual
compliance parameter. This parameter determines whether she takes up rapid
tests. Positive tests and symptoms lead most individuals to reduce their contacts.

Modelling a population of agents according to actual demographic characteristics means
that we can use a wide array of data to identify and calibrate the model's many
parameters.\footnote{See Supplementary Material~\ref{sec:materials_and_methods} for a
    complete description.} Contact diaries yield pre-pandemic distributions of contacts for
different contact types and their assortativity by age group. Mobility data is used to
model the evolution of work contacts. School and daycare policies can be incorporated
directly from official directives. Administrative records on the number of tests,
vaccinations by age and region, and the prevalence of virus strains are generally
available. Surveys may ask about test offers, propensities to take them up, and past
tests. Other studies' estimates of the seasonality of infections can be incorporated
directly. The remaining parameters---most notably, these include infection probabilities
by contact network and the effects of some NPIs, see Supplementary
Material~\ref{subsec:estimated_params}---will be chosen numerically so that the model
matches features of the data \citep[see][for the general method]{McFadden1989}.

@@

Until mid-March, there is no visible difference between the different scenarios.
Seasonality hardly changes, and only few vaccinations and rapid tests were administered.
Even thereafter, the effect of the vaccination campaign is surprisingly small.
The Shapley value decomposition shows that vaccinations contribute 16\% to the
cumulative difference between scenarios. Reasons for the low share are the slow
start---it took until March~24th until 10\% of the population had received their first
vaccination, the 20\% mark was reached on April 19th---and the focus on older
individuals. By the end of our study period, when first-dose
vaccination rates reached 43\% of the population, the numbers of new cases would have
started to decline. However, the vaccination campaign reached its goal of protecting
risk groups.

@@

A similar-sized effect---42\% in the decomposition---comes from rapid testing. Here, it
is crucial to differentiate between recorded cases and actual cases. Additional testing
means that additional infections will be recorded which would otherwise remain
undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this effect is large and
may persist for some time. Until late April, recorded cases are higher in the scenario
with rapid testing alone when compared to the setting where none of the three mechanisms
are turned on. The effect on total cases, however, is visible immediately in
Figure~\ref{fig:2021_scenarios_newly_infected}. Despite the fact that only 10\% of the
population performed weekly rapid tests in March on average, new infections on April~1
would have been reduced by 53\% relative to the scenario without vaccinations, rapid
tests, or seasonality.

So why is rapid testing so effective? In order to shed more light on this question,
Figure~\ref{fig:2021_scenarios_decomposition_tests} decomposes the difference in the
scenario without rapid tests and the main specification into the three channels for
rapid tests. Tests at schools have the smallest effect, which is largely explained by
schools not operating at full capacity during our period of study and the relatively
small number of students.\footnote{18\% of our population are in the education sector
    (pupils, teachers, etc.); 46\% are workers outside the education sector.} Almost 40\%
come from tests at the workplace. Despite the fact that rapid tests for private reasons
are phased in only in mid-March, they make up for more than half of the total effect.
The reason lies in the fact that a substantial share of these tests is driven by an
elevated probability to carry the virus, i.e., showing symptoms of CoViD-19 or following
up on a positive test of a household member. The latter is essentially a form of contact
tracing, which has been shown to be very effective \citep{Contreras2021,
    Fetzer2021,Kretzschmar2020}. A deeper analysis in Supplementary
Material~\ref{subsec:appendix_scenarios} shows that the same amount of rapid tests
administered randomly in the population would not have been nearly as effective and that
mandatory tests at schools almost offset the increased infection risk of opening
schools (see Figure~\ref{fig:school_scenarios}).

>>>>

At the core of our agent-based model \citep[][we review more literature in Supplementary
Material~\ref{sec:literature_review}]{Aleta2020,Hinch2020} are physical contacts between
heterogeneous agents (Figure~\ref{fig:model_contacts_infections}). Each contact between
infectious and susceptible individuals bears the risk of transmitting the virus.
Contacts occur in up to four networks: Within the household, at
work, at school, or in other settings (leisure activities, grocery shopping, medical
appointments, etc.). Some contacts recur regularly, others occur at random. Empirical
applications can take the population and household structure from census data and the
network-specific frequencies of contacts from diary data measuring contacts before the
pandemic \citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, meeting frequencies
depend on age and geographical location (see Supplementary
Material~\ref{subsec:assortativity}).

The four contact networks are chosen so that the most common NPIs can be modeled in
great detail. NPIs affect the number of contacts or the risk of transmitting the disease
upon having physical contact. The effect of different NPIs will generally vary across
contact types. See Section~\ref{sec:policies} for details on how we model NPIs.
 In any setting, mask mandates or air filtes would reduce the probability
of infection \citep{Lessler2021, Cheng2021}.

@@

In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age. A
possible infection progresses as shown in Figure~\ref{fig:model_disease_progression}. We
differentiate between an initial period of infection without being infectious or showing
symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring
intensive care, and recovery or death \citep[similar to][]{Aleta2020}. The calibration
of the age dependent probabilities of transitioning between these states  is described
in Supplementary Material~\ref{sub:data_course_of_disease}). Conditional on the type of
contact, infectiousness is independent of age \citep{Jones2021}.

The model includes several other crucial features. New virus strains with different
infectiousness profiles may appear. Vaccines may become available. During the vaccine
roll-out, priority may depend on age and occupation; vaccine hesitancy is modelled by
some individuals refusing vaccination offers. With some probability, vaccinated agents
become immune and do not transmit the virus \citep{Hunter2021, LevineTiefenbrun2021,
Petter2021, Pritchard2021}.

We include two types of tests. Polymerase chain reaction (PCR) tests reveal with
certainty whether an individual is infected or not; They require at least one day to be
processed and are scarce. In contrast, rapid antigen tests yield immediate results.
Specificity and sensitivity of
these tests is set according to data analyzed in \cite{Bruemmer2021, Smith2021};
sensitivity depends on the timing of the test relative to the onset of infectiousness.
Figure~\ref{fig:model_rapid_tests} shows our model for rapid test demand. Schools may
require staff and students to be tested regularly. Rapid tests may be offered by
employers to on-site workers. Individuals may demand tests for private reasons, which
include having plans to meet other people, showing symptoms of CoViD-19, and a household
member having tested positively for the virus. We endow each agent with an individual
compliance parameter. This parameter determines whether she takes up rapid
tests. Positive tests and symptoms lead most individuals to reduce their contacts.

Modelling a population of agents according to actual demographic characteristics means
that we can use a wide array of data to identify and calibrate the model's many
parameters.\footnote{See Supplementary Material~\ref{sec:materials_and_methods} for a
    complete description.}
Examples are contact diaries, administrative records on case numbers, virus straints,
tests and vaccines as well as surveys on private rapid test demand.
Other studies' estimates of the seasonality of infections can be incorporated
directly. The remaining parameters---most notably, the infection probabilities
by contact network and the effects of some NPIs, see Supplementary
Material~\ref{subsec:estimated_params}---will be chosen numerically so that the model
matches features of the data \citep[see][for the general method]{McFadden1989}.

@@

Until mid-March, there is no visible difference between the different scenarios.
Seasonality hardly changes, and only few vaccinations and rapid tests were administered.
Even thereafter, the effect of the vaccination campaign is surprisingly small.
The Shapley value decomposition shows that vaccinations contribute 16\% to the
cumulative difference between scenarios. Reasons for the low share are the slow
start---it took until March~24th until 10\% of the population had received their first
vaccination, the 20\% mark was reached on April 19th---and the focus on older
individuals. By the end of our study period, when first-dose
vaccination rates reached 43\% of the population, the numbers of new cases would have
started to decline.

@@

A similar-sized effect---42\% in the decomposition---comes from rapid testing. Here, it
is crucial to differentiate between recorded cases and actual cases. Additional testing
means that additional infections will be recorded which would otherwise remain
undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this effect is large and
may persist for some time. Until late April, recorded cases are higher in the scenario
with rapid testing alone when compared to the setting where none of the three mechanisms
are turned on. The effect on total cases, however, is visible immediately in
Figure~\ref{fig:2021_scenarios_newly_infected}.

% Despite the fact that only 10\% of the
% population performed weekly rapid tests in March on average, new infections on April~1
% would have been reduced by 53\% relative to the scenario without vaccinations, rapid
% tests, or seasonality.

So why is rapid testing so effective? In order to shed more light on this question,
Figure~\ref{fig:2021_scenarios_decomposition_tests} decomposes the difference in the
scenario without rapid tests and the main specification into the three channels for
rapid tests. Tests at schools have the smallest effect, which is largely explained by
schools not operating at full capacity during our period of study and the relatively
small number of students.\footnote{18\% of our population are in the education sector
    (pupils, teachers, etc.); 46\% are workers outside the education sector.} Almost 40\%
come from tests at the workplace. Despite the fact that rapid tests for private reasons
are phased in only in mid-March, they make up for more than half of the total effect.
The reason lies in the fact that a substantial share of these tests is driven by an
elevated probability to carry the virus, i.e., showing symptoms of CoViD-19 or following
up on a positive test of a household member. This is essentially a form of contact
tracing, which has been shown to be very effective \citep{Contreras2021,
    Fetzer2021,Kretzschmar2020}. Indeed, a deeper analysis in Supplementary
Material~\ref{subsec:appendix_scenarios} shows that the same amount of rapid tests
administered randomly in the population would not have been nearly as effective and that
mandatory tests at schools almost offset the increased infection risk of opening
schools (see Figure~\ref{fig:school_scenarios}).
